---
title: "PA1_template.Rmd"
author: "Venkat"
date: "7/26/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ReproducibleResearch: Week2 Assignment for Peer Review

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(mice)
```

## Step1: Load and Read Data.
```{r, echo=TRUE, results='show', message=FALSE, warning=FALSE}
#download file directly from online and extract it
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode = "wb")
unzip("activity.zip")

#Read the data and check the contents
activity_data <- read.csv("activity.csv", header = TRUE)
head(activity_data)
```

## Step 2: Histogram of total steps/day
```{r, echo=TRUE, results='show', message=FALSE, warning=FALSE}
# aggregate the 5 min data into day level
stepsbydate <- activity_data %>% select(date,steps) %>% group_by(date) %>% summarize(totalsteps = sum(steps)) %>%na.omit()
#plot the histogram
hist(stepsbydate$totalsteps, xlab="Total Daily Steps", ylab="Freq", main="Histogram of Steps by Day", breaks = 15)
```

<br/>

## Step 3: Mean and Median of steps/day
```{r}
stepsbydateMean <- mean(stepsbydate$totalsteps)
stepsbydateMedian <- median(stepsbydate$totalsteps)

```
* Mean: `r stepsbydateMean`
* Median: `r stepsbydateMedian`


## Step 4: Time series Plot of Avg Steps taken per day
```{r}
databyinterval <- activity_data %>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(totalsteps = mean(steps))
## Now plot the summary of steps
ggplot(databyinterval, aes(x=interval, y=totalsteps))+geom_line()+labs(title="Time Series Plot", y="Avg # of steps taken", x="interval - 5 min")
```

## Step 5: The 5-minute interval where average contains max #of steps
```{r}
databyinterval[which(databyinterval$totalsteps == max(databyinterval$totalsteps)),]
```
## Step 6: Code to describe and show the strategy for imputing the missing data
```{r}
count_of_missing_values = length(which(is.na(activity_data$steps)))
summary(activity_data)
```
* missing data rows: `r count_of_missing_values`
<br/>
Use impute function with mean and fill steps for the days they are missing.

```{r}
library(dplyr)
replacewithmean <- function(num) replace(num, is.na(num),mean(num, na.rm = TRUE))
activitydata_nomissing <- activity_data %>% group_by(interval) %>% mutate(steps = replacewithmean(steps))
head(activitydata_nomissing)
```

```{r}
new_activity_data = as.data.frame(activitydata_nomissing)
head(new_activity_data)

```

```{r}
count_of_missing_values2 = length(which(is.na(new_activity_data$steps)))

```
* Number of missing values(steps): `r count_of_missing_values2`

```{r}
summary(new_activity_data)
```
## Step 7: Histogram of Total steps taken for each day after missing values are imputed

For the histogram sum up the steps for each day
```{r}
day_summary <- aggregate(new_activity_data$steps, by=list(new_activity_data$date), sum)

names(day_summary)[1] = "date"
names(day_summary)[2] = "totalsteps"
head(day_summary)

```

```{r}
hist(day_summary$totalsteps, xlab="Total Daily Steps",ylab="Freq" ,main="Histogram of Steps by Day after imputatiob", breaks = 15)
```


## Step 8: Panel split comparing the average number of steps taken per 5min interval across week days and weekends

```{r}
new_activity_data$weekend_flag <- ifelse(weekdays(as.Date(new_activity_data$date)) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),"Weekday","Weekend")
head(new_activity_data)
```

```{r}
new_activity_data <- (new_activity_data %>% group_by(interval,weekend_flag) %>% summarise(Mean= mean(steps)))

ggplot(new_activity_data, mapping = aes(x=interval, y=Mean)) + geom_line()+
facet_grid(weekend_flag ~.) +xlab("Interval") +ylab("Mean of steps") + ggtitle("comparision of steps for each interval")
```

## Step 9: All of R code needed to reproduce the results in the report.